---
title: "NHS_R_purrr"
output: html_document
date: '2022-06-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
# Creating an argument for rolling dice
# This is an example of a function factory

dice <- function(sides) {
  function(rolls) {
    sample(1:sides, rolls, TRUE)
  }
}

six_sided <- dice(6)

six_sided(2)
```


# Named vectors

```{r}
named_vector <- c(a = 1, b = 2, c = 3)

# Now you can access a value using the name 

named_vector[["a"]]

# you can also set names using the set_names argument in purrr 

letters[1:3] %>% set_names()
```

# Dataframes

A dataframe is just a named list, but every tiem in the list is a vector. The names of the items in the list are the colummn name. Each item in a dataframe is a vector 

# Iteration 

```{r}
my_values <-  c(1, 4, 5, 3, 2)

my_fun <- function(x) {
  3 * x + 2
}

# because R is vectorised through out, you can simply run: 

my_fun(my_values)

# and it will iterate over every part of th the list. 
# However not all functions work like this and sometimes you'll have to write a loop
# Not all functions are vectorised
```

```{r}
# doing the same thing but with a loop

# First, we set up a vector that is the same length as my_values so that it can accept the results 

results <- numeric(length(my_values))

# Then, we set a loop to iterate over the values 

for (i in 1:length(my_values)) {
  results[[i]] <- my_fun(my_values[[i]])
}

results
```

The problems with writing a loop include:
* Forgetting to the initialise the results vector
* Incorrectly setting up the results vector 
* Mess up the indexing, for example writing `my_values[[1]]` instead of `my_values[[i]]`

# Map 

However the function `map` can replace loops. It takes the input that you wish to iterate over, and a function that you want to evaluate each item from the input against 

```{r}
results <-  map(my_values, my_fun)

results
# But of course, this returns a list instead of a vector
# Which is why if you do something like this (below), it will throw out an error

mean(results)

# The solution is then to use a function to tell what map to return i.e. 

results <- map_dbl(my_values, my_fun)

results 

# Of course, one can also use an anonymous function 

map_dbl(my_values, ~ 3 * .x + 2)

# Use the formula syntax if it's really simple 
```

```{r}
# If your function has more than one argument, these can be passed in the "..." of the map argument

my_fun <- function(x, y) {
  3 * x + y
}

map_dbl(my_values, my_fun, y = 2)
```

# Using map with dataframes

```{r}
# This produces far more valuable information than 
map_chr(iris, class)
# This
class(iris)
```

```{r}
# This first collects the unique values in each column, then it prints out the number of unique columns 
iris %>% 
  map(unique) %>% 
  map_dbl(length)

# We're able to chain functions together to obtain a more complex result --
# I think this would be particularly useful for this like is.na
```

You can also use map with a mutate statement. This is particularly useful if you want to create a column of values from another column, but that function is not vectorised. 


```{r}
files <- dir("./nhs_webinars/fp_with_purrr/ae_attendances", 
             "^\\d{4}-\\d{2}-\\d{2}\\.csv$", 
             full.names = TRUE)

head(files)

# If you were to run read_csv with `files`, it would produce a very weird result 
# Instead, we should use map to iterate over all of these files to call the read_csv function 

map(files, read_csv, col_types = "ccddd")

# However, it would be even better to return one single dataframe by binding all rows

map_dfr(files, read_csv, col_types = "ccddd")

# Additionally, `map_dfr` has a really neat trick up its sleeves, which is the `.id` argument, 
# which takes the names of each item in the vector and adds it to a column with the name that you give it. 
# However, we first need to create a named_vector

ae_attendances <- files %>% 
  set_names() %>% 
  map_dfr(read_csv, col_type = "ccddd", .id = "filename") %>% 
  mutate(period = str_extract(filename, 
                              r"(\d{4}-\d{2}-\d{2}(?=\.csv$))") %>% 
           lubridate::ymd())
```

# More variants of map

```{r}
# `map2` takes two vectors as arguments, .x and .y and a binary function .f
map2_dbl(1:3, 4:6, ~ .x * .y)

# `pmap` is a generalisation of map that works over as many many vectors as you have 

list(1:3, 4:6, 7:9) %>% 
  pmap_dbl(function(x, y, z) x * y + z)

# Unless you have a named vector, then purrr will match the list items to the function arguments 
list(a = 1:3, b = 4:6, c = 7:9) %>% 
  pmap_dbl(function(c, b, a) a * b + c)
```

# Walk 

Walk is like map, except it is designed for functions that you run solely for the side effects. 
Instead of returning the results of the initial function, `walk` returns the input vector. 
`Walk` is useful for operations like saving plots to disk. 

# Demo 

```{r}
ae_attendances %>% 
  filter(str_starts(org_code, "R")) %>% 
  group_by(org_code, period) %>% 
  summarise_at("attendances", sum) %>% 
  # He thinks it's a lot neater than: 
  # summarise(attendances = sum(attendances))
  nest() %>% 
  #if you nest, it will nest by how it is grouped
  mutate(has_3_years = map_dbl(data, nrow) == 36)
  # This checks if each tibble has 36 rows, which is equivalent to 3 years of data
  # you can also just filter it outright with filter(map_dbl(data, nrow) == 36)

# If we were just to use nrow on its own, it would throw up an error of 
# Result must have length 1, not 0

# Creating a plot function
plot_fn <- function(org_code, data) {
  data %>% 
    ggplot(aes(x = period, y = attendances)) + 
    geom_line() + 
    geom_point() + 
    labs(title = org_code)
}

# Creating a new folder to store the plots 
dir.create("./nhs_webinars/fp_with_purrr/ae_plots")

ae_attendances %>% 
  filter(str_starts(org_code, "R")) %>% 
  group_by(org_code, period) %>% 
  summarise_at("attendances", sum) %>% 
  nest() %>% 
  filter(map_dbl(data, nrow) == 36) %>% 
  # Creating a column in a dataframe for the plot? Let's see how this works 
  mutate(plot = map2(org_code, data, plot_fn)) %>% 
  mutate(filename = paste0("ae_plot/", org_code, ".png")) %>% 
  ungroup() %>% 
  select(-org_code, -data) %>% 
  head(10) %>% 
  # I have no idea why this is not working 
  pwalk(ggsave)
  pwalk(ggsave(paste0(filename), plot = plot, width = 11, height = 8.5, units = "in"))
  
ggsave()
  filename = paste0(filename, ".pdf"), plot = temp_chart, width = 11, height = 8.5, units = "in"
# This prints out what the data grouping
groups(.Last.value)
  
```

