---
title: "Untitled"
output: html_document
date: '2022-05-07'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(skimr)
library(janitor)

theme_set(theme_light())

baseball_wins <- read_csv("https://raw.githubusercontent.com/thebioengineer/TidyX/master/TidyTuesday_Explained/023-Mailbag_and_Expected_Wins/Pythagorean%20Wins.csv") %>% 
  clean_names()

```


```{r}
glimpse(baseball_wins)

baseball_wins %>% skim()
```

```{r}
baseball_ratio <- baseball_wins %>%  
  mutate(scoring_ratio = runs_scored / runs_allowed, 
         win_pct = wins / (wins + losses))
```

Calculating the Pythagorean wins percentage for each team 

$PyrhWinPct = Scoring Ratio^i / (1 + Scoring Ratio^i)$

The exponent `i` will initially `3` but we will optimise this with our `for` loop. 

To change odds into a percentage, you divide it by 1 + odds 

```{r}
i <- 3

br_pythag <- baseball_ratio %>%
  mutate(win_pct_scale = scoring_ratio^i / (1 + scoring_ratio^i))


  
```

```{r}
baseball_error <- br_pythag %>% 
  mutate(error = win_pct - win_pct_scale)

baseball_error %>% 
  ggplot(aes(x = error)) + 
  geom_histogram(alpha = .5) + 
  geom_vline(xintercept = 0, size = 1.2, colour = "red", linetype = "dashed") + 
  labs(title = "Error between actual win% and pythagorean win%")

```

```{r}
baseball_error %>% 
  ggplot() + 
  geom_histogram(aes(x = win_pct), fill = "light blue", alpha = .6) + 
  geom_histogram(aes(x = win_pct_scale), fill = "light green", alpha = .5) + 
  annotate("text", x = .3, y = 70, label = "Actual win%", colour = "light blue", size = 4) + 
  annotate("text", x = .3, y = 60, label = "Pyathagorean win%", colour = "light green", size = 4) + 
  ggtitle("Actual win% and Pythagorean win%")
```

```{r}
mae <- mean(abs(baseball_error$error))
```

### how to optimise exponent, using for loops 

```{r}
i_options <- seq(from = .5, to = 4, by = .1)

# creating a df for the for loop 
# pre-allocating an empty df will speed up for loops a lot

mae_results <- data.frame("exp" = i_options, "results" = NA)

df2 <-  baseball_ratio[, 3:6]

```

```{r}
for(i in 1:length(i_options)){
  
  win_pct <-  df2$wins / (df2$wins + df2$losses)
  
  pred_win_pct <- (df2$runs_scored / df2$runs_allowed)^i_options[i] / ((1 + df2$runs_scored / df2$runs_allowed)^i_options[1])
  
  mae_results[i, 2] <- mean(abs(win_pct - pred_win_pct))
}
```

### optimising with functions and tidyverse

```{r}

# abstractifying the calculation 

pythagorean_wins <-  function(scored, allowed, i){
  
  p_win_pct <- (scored / allowed)^i / (1 + (scored / allowed)^i)
  
  return(p_win_pct)
}
```

```{r}
for( i in 1:length(i_options)){
  
  baseball_ratio_options <- baseball_ratio %>%  
    mutate(pred_win_pct = pythagorean_wins(
      scored = runs_scored, 
      allowed = runs_allowed, 
      i = i_options[[i]]
    )) %>% 
    select(win_pct, pred_win_pct)
  
  mae_results[i, 2] <- mean(abs(baseball_ratio_options[[1]] - baseball_ratio_options[[2]]))
}
```

### plottinf the exponent as a line

```{r}
mae_results %>%  
  ggplot(aes(x = exp, y = results)) + 
  geom_line() + 
  labs(x = "Exponent", y = "Mean absolute error", 
       title = "Mean absolute erro for different exponents in Pyathagorean wins formula")
```


```{r}
min(mae_results$results)

mae_results %>% filter(results == min(results))

```

### optimising with the tidyverse approach 

```{r}
mae_tibble <- seq(from = .5, to = 4, by = .1) %>% 
  
  map_dfr(function(i){
    
    baseball_ratio_options <-  baseball_ratio %>% 
      mutate(pred_win_pct = pythagorean_wins(
        scored = runs_scored, 
        allowed = runs_allowed, 
        i = i
      )) %>%  
      select(win_pct, pred_win_pct)
    
    tibble(
      exp = i,  
      results = mean(abs(baseball_ratio_options[[1]] - baseball_ratio_options[[2]]))
    )
  })

mae_tibble %>% 
  filter(results == min(results))
```

